---
title: "Testing Code Blocks"
author: Brad Cannell
date: 2025-04-18
date-modified: 2025-04-18
notes: 
  - Use a Quarto document instead of a plain MD document for the outline in 
    RStudio.
details:
  - The prompt below assumes that I'm in the EWB project on ChatGPT, and 
    therefore, don't need to explain what EWB is or pass along all of the formatting rules
    contained in the style guide.
  - I got the best results with ChatGPT 04-mini-high.
  - So far, I can only get this to work if I give it instructions one step at a time.
  - Additionally, the SCTs almost never work yet.
---

# Attach Coding Lesson Authoring Guide

I'm attaching a Quarto markdown document titled, "Coding Lesson Authoring Guide.qmd". Please use it as a resource in this chat.

-----

# Initial prompt

Later, will attach a coding exercise from the "Introduction to R Programming and Data Management" I teach at a university as a Quarto markdown document. Please help adapt the Quarto markdown document for use as a coding exercise on the Epi-Workbench platform. The Quarto markdown document contains one or more code blocks, which we will refer to as "exercise code blocks."

Please execute the following steps each time you encounter an exercise code block:

## Step 1: Create a solution code block

- Create a solution code block below the exercise code block - separated by a single line break.
- Whenever you see a code block that begins with exactly `{r}` and ends with a line that’s exactly "```", copy that entire block—including both back‑tick lines—verbatim.
- Detect the start of an exercise block with a regex like ^```\\{r\\}\\s*$ and its end with ^```\\s*$. Everything between (and including) those two lines is the exercise block.
- Do not change anything inside the exercise block. Only add new code fences below it.
- The solution code block should be identical to the exercise code block except that it will use the `type=solution` code block option in the code block header. 

### Example

Attached Quarto markdown document contains:

````
```{r}
x <- 15
```
````

The solution block should be added like this:

````
```{r}
x <- 15
```

```{r, type=solution}
x <- 15
```
````

## Step 2: Create a hidden code block for local testing

- Create a hidden code block below the solution code block - separated by a single line break.
- The hidden code block will use the `type=hide` code block option in the code block header.
- The contents of the hidden code block should follow this pattern:

```{r, type=hide}
# Hidden Code Block
# Run local tests that are hidden from students.
# No error = no problem.

# Solution code
solution <- [Add solution code here]

# Simulate a student submission
submission <- [Add submission code here]

# Setup the exercise state
setup_state(
  sol_code = solution,
  stu_code = submission
)

# Submission correctness test (SCT)
[Add submission correctness test here]
```

Additionally:
- Please replace `[Add solution code here]` with the code from the solution code block.
  - The solution code must be wrapped in single or double quotes. For example, `solution <- "x <- 15"` instead of `solution <- x <- 15`.
- Please replace `[Add submission code here]` with the code from the solution code block.
  - The submission code must be wrapped in single or double quotes. For example, `submission <- "x <- 15"` instead of `submission <- x <- 15`.
- Please replace `[Add submission correctness test here]` with a submission correctness test (SCT) from the `testwhat` package.
  - The documentation for the `testwhat` package is available at: https://datacamp.github.io/testwhat/
  - All SCTs should begin with `ex() |>` verbatim.

## Step 3: Create a test code block

- Create a test code block below the hidden code block - separated by a single line break.
- The test code block will use the `type=test` code block option in the code block header.
- The test code block should contain the SCT from the hidden code block immediately above.

### Example

The hidden block contains:

````
```{r, type=hide}
# Hidden Code Block
# Run local tests that are hidden from students.
# No error = no problem.

# Solution code
solution <- "library(dplyr)"

# Simulate a student submission
submission <- "library(dplyr)"

# Setup the exercise state
setup_state(
  sol_code = solution,
  stu_code = submission
)

# Submission correctness test (SCT)
ex() %>% check_library("dplyr")
```
````

The test block should be added like this: 

````
```{r, type=hide}
# Hidden Code Block
# Run local tests that are hidden from students.
# No error = no problem.

# Solution code
solution <- "library(dplyr)"

# Simulate a student submission
submission <- "library(dplyr)"

# Setup the exercise state
setup_state(
  sol_code = solution,
  stu_code = submission
)

# Submission correctness test (SCT)
ex() %>% check_library("dplyr")
```

```{r, type=test}
ex() %>% check_library("dplyr")
```
````

## Step 4: Create hints

- Write a set of hints for learners below the test code block - separated by a single line break and using the appropriate hint syntax.
- Please write 2-5 hints that are progressively more helpful to learners.
- The final hint should contain the complete solution.
- Each exercise code block is worth 100 knowledge points (KP).
 - Each hint should deduct 100 / [total number of hints] KP.
 - Please calculate the KP deduction for each hint and add it to the hint using the `[POINTS=X]` syntax.
- Please write hints using the syntax from "Coding Lesson Authoring Guide.qmd".

### Example

Incorrect hint syntax:

````
> [!hint] **Hint 1 [POINTS=-25]**
> Make sure the object `x` exists before trying to print it.

> [!hint] **Hint 2 [POINTS=-25]**
> Remember that `print()` is the function used to display objects in R.

> [!hint] **Hint 3 [POINTS=-25]**
> Inside `print()`, include the object name: `print(x)`.

> [!hint] **Hint 4 [POINTS=-25]**
> **Solution:** `print(x)`
````

Correct hint syntax: 

````
<!-- HINT:
[POINTS=-25]
- Make sure the object `x` exists before trying to print it.

[POINTS=-25]
- Remember that `print()` is the function used to display objects in R.

[POINTS=-25]
- Inside `print()`, include the object name: `print(x)`.

[TITLE=Solution, POINTS=-25]
- `print(x)`
-->
````

## Other instructions
    
- Please don't change anything else about the Quarto markdown document.
- When you have completed the steps above, please create an updated Quarto markdown document that I can download.


