---
title: "Experimenting with a New Multi-step CST Creation Workflow that Uses testthat Expectations"
id: multi_step_cst_creation
author: brad_cannell
date: 2026-01-29
date-modified: 2026-01-29
audience: course_authors
category: lesson_development
tags:
  - r
  - csts
  - test_that
  - unit_testing
format:
  html:
    toc: true
    embed-resources: true
execute:
  eval: false
---

# Overview

-   I'm experimenting with a new multi-step CST creation workflow that uses `testthat` expectation functions.
-   New lessons: Take multiple passes to complete CSTs. Pass in the entire lesson file.
    1.  User: Add scaffolding, solution block, and instructions to the lesson Quarto file.
    2.  AI: For each exercise, write example incorrect responses. Return them in hidden code blocks.
    3.  AI: For each incorrect response, write one or more `testthat` tests using `expect` functions.
        -   Maybe create a vector store or RAG with `testthat` documentation and EWB SOPs.
    4.  AI: For each exercise, write hints.
        -   Again, use a vector store or RAG with EWB SOPs.
-   2026-01-29: For now, I'm just having this conversation in the interactive app for quick iteration. I'm not passing a prompt.

# Prerequisites

-   Create a new virtual environment (venv).
-   Install Packages:
    -   pip install keyring
    -   pip install openai
    -   pip install jupyter
    -   pip install -U chatlas
-   pip freeze \> requirements.txt

# Load Packages

```{python}
from datetime import date
from pathlib import Path
from openai import OpenAI
import chatlas as ctl
import keyring
import json
```

# File Paths and Environment Variables

-   Set today's date.
-   Set the course name.
-   Set the path to the lesson folder.
-   Set the path to the lesson file.
-   Get the content of the lesson file.
-   Set the path to the lesson revision prompt.
-   Set the model to inject into the API calls. See available models at: https://platform.openai.com/docs/models.

They are passed to the AI assistant below.

```{python}
# OpenAI Models: https://platform.openai.com/docs/models
# Anthropic Models: https://platform.claude.com/docs/en/about-claude/models/overview
model = "gpt-5-nano-2025-08-07"
today = date.today().isoformat()

# Conversation
conversation_folder_path = Path("conversations/CSTs/Multi-step CST Creation/")

# Prompt paths
path_Template_Prompt_CST_01_Instructions = Path('/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/prompts/Courses/Template Prompt - CST 01 Instructions.md')
path_Template_Prompt_CST_02_CST = Path('/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/prompts/Courses/Template Prompt - CST 02 CST.md')

# Prompt content
# I'm going to read in the prompts above in case I want to pass them to the model
# when explaining our current process below.
Template_Prompt_CST_01_Instructions = path_Template_Prompt_CST_01_Instructions.read_text(encoding='utf-8')
Template_Prompt_CST_02_CST = path_Template_Prompt_CST_02_CST.read_text(encoding='utf-8')
```

# Load helper functions

Run in terminal: 
- `uv add --editable "/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/py_scripts"`

```{python}
# Load functions from py_scripts/ewb_py_scripts
from ewb_py_scripts import (
    prompt_extract_system_prompt,
    prompt_extract_user_message,
    prompt_fill_placeholders,
)
```

# Expose API

This assumes you've already created an API key on [openai's website](https://platform.openai.com/api-keys), and that you saved the API key in keyring using `keyring.set_password("API Name", "User", "API Key")`.

```{python}
# Retrieve the password from the system keyring
openai_api = keyring.get_password("ewb-course-development", "brad")

if openai_api is not None:
    # Set the retrieved value as a temporary environment variable
    os.environ['OPENAI_API_KEY'] = openai_api
    print("Environment variable 'OPENAI_API_KEY' set successfully within the script.")
else:
    print("Could not retrieve password from keyring.")
```

# Create Chat Object

```{python}
chat = ctl.ChatOpenAI(
    model = model
)
```

# Quick Iteration Chat

-   2026-01-29: For now, I'm just having this conversation in the interactive app for quick iteration. I'm not passing a prompt.
    -   Chat in interactive mode - `chat.app()`.

## Conversation starter (pass everything below to the chat app)

I am an epidemiologist, gerontologist, data scientist, and professor at Texas Christian University. I am also the co-founder and owner of Epi-Workbench, LLC - a small business developing a web-based educational platform.

I am creating a web-based business called "Epi-Workbench" (EWB). EWB is an online platform offering interactive courses in epidemiology, data science, and artificial intelligence specifically designed for people working in public health and other health professions and, to a lesser extent, members of the general public who are interested in learning more about health research. With beginner to advanced-level courses, EWB combines real-person instruction with hands-on coding exercises and integrated AI tools for a comprehensive learning experience. Learners can develop skills in statistical programming, applied statistical analyses, and scientific communication—all from the convenience of their browser, with no installation required.

The coding exercises embedded in EWB lessons are typically made up of a sequence of four code blocks, each serving a specific purpose:

1.  Exercise Code Blocks ({r}): Used for executable code where learners write, edit, and submit their solutions. The only blocks that are visible to learners on the EWB platform.
2.  Solution Code Blocks ({r,type=solution}): Used to document the correct solution for an exercise. Not visible to learners.
3.  Hidden Code Blocks ({r,type=hide}):Used during development to run or test code locally, including Code Submission Tests (CSTs). Not visible to learners.
4.  Test Code Blocks ({r,type=test}): Used to define CSTs that evaluate learner submissions. Not visible to learners.

### Test code blocks

-   Test code blocks contain **Code Submission Tests (CSTs)** used by the EWB platform to evaluate whether learners submitted correct code.\
-   The platform uses the CSTs in the test code block to check the correctness of the exercise code block that precedes it.\
-   Test code blocks must **follow** an exercise code block and are automatically associated with the closest preceding exercise block.\
-   Use the `type=test` option in the code block header to designate a test code block.\
-   You can also insert a test code block directly into the active Quarto lesson file using the **Insert Test Code Block** RStudio Addin (available in the [EWBTemplates](https://github.com/epi-workbench/EWBTemplates) package).

#### Example coding exercise containing all four code block types

Instuctions to learners: Before we dive into a discussion about descriptive analysis, let's simulate some data to practice with below. Start by completing the code below to load the `dplyr` package.

```` markdown
```{r}
# Load the dplyr package
____
```

```{r, type=solution}
# Load the dplyr package
library(dplyr, warn.conflicts = FALSE)
```

```{r, type=hide}
# Hidden Block for Local Testing Only
# -----------------------------------------------------------------------------

# Setup the simulated learner environment
learner_env <- new.env()
rm(list = ls(envir = learner_env), envir = learner_env)

# Simulated learner submissions
correct <- 'library(dplyr, warn.conflicts = FALSE)'
correct_without_warn_conflicts <- 'library(dplyr)'
correct_quoted_pkg <- 'library("dplyr")'
wrong_no_change <- '____'
wrong_missing_pkg <- 'library(ggplot2)'
wrong_fn <- 'paste(dplyr)'
wrong_require <- 'require(dplyr)' # Not best practice, but it works

# Code for detaching packages during interactive testing
# detach("package:dplyr", unload = TRUE)

# Set the active submission
learner_code <- correct

# Gracefully evaluate code (prevents early error from stopping tests)
try(eval(parse(text = learner_code), envir = learner_env), silent = TRUE)
```

```{r, type=test}
#| label: cst-lesson-01

# Check that `____` was replaced with something
test_that("All blanks `____` were replaced", {
  expect_false(
    # Match two or more consecutive underscores (regex)
    grepl("__+", learner_code)
    info = "It looks like your submission still contains `____`. Please replace `____` to complete the code."
  )
})

# Check that packages were loaded
test_that("Did you load the required package?", {
  pkg <- "package_name" # Replace package_name
  loaded_pkgs <- eval(.packages(), envir = learner_env)
  expect_true(
    tolower(pkg) %in% tolower(loaded_pkgs),
    info = glue::glue(
      "Did you correctly load the `{pkg}` package using the `library()` function?"
    )
  )
})
```
````

### Current process

When developing courses, our current process is:

1.  We write and refine the lesson text in a Quarto file. Initially, solution code blocks are the only type of coding exercise blocks in the Quarto lesson file.

2.  We create an exercise code block with code scaffolding.

3.  We write instructions for completing each coding exercise.

4.  We open a ChatGPT chat.

5.  We pass each set of instructions, exercise code block, and solution block to ChatGPT using the prompt titled, "Template Prompt - CST 01 Instructions" (provided below).

6.  ChatGPT returns suggestions for hidden code blocks and test code blocks containing CSTs.

#### Template Prompt - CST 01 Instructions

'---\\ntitle: Template Prompt for EWB Code Submission Tests (CSTs) - Step 1 - Instructions\\nid: template_prompt_cst_instructions\\ntask: code_review\\naudience: academic\\nauthor: brad_cannell\\ncreated: 2026-01-17\\nlast_updated: 2026-01-17\\nproject: null\\ndomain: lesson_development\\ntags:\\n  - ewb\\n  - csts\\n---\\n\\n# Overview (optional)\\n\\nThis prompt is for getting assistance with writing code submission tests for EWB lessons.\\n\\n- I want to provide:\\n  - The instructions learners will read immediately before the coding exercise.\\n  - The scaffolded code given to learners.\\n  - The solution code. What the scaffolded code should look like when learners correctly complete the coding exercise.\\n\\n- I want the AI assistant to return:\\n  - Realistic incorrect submissions that learners might submit.\\n  - Code submission tests written with code from the testthat package.\\n\\n\\n# SYSTEM PROMPT\\n\\nYou are an R developer and instructional designer with deep expertise in writing testthat-based Code Submission Tests (CSTs) for educational platforms.\\n\\n---\\n\\n# User Message\\n\\n## Context\\n\\nI am an epidemiologist, gerontologist, data scientist, and professor at Texas Christian University. I am also the co-founder and owner of Epi-Workbench, LLC - a small business developing a web-based educational platform.\\n\\nI am creating a web-based business called "Epi-Workbench" (EWB). EWB is an online platform offering interactive courses in epidemiology, data science, and artificial intelligence specifically designed for people working in public health and other health professions and, to a lesser extent, members of the general public who are interested in learning more about health research. With beginner to advanced-level courses, EWB combines real-person instruction with hands-on coding exercises and integrated AI tools for a comprehensive learning experience. Learners can develop skills in statistical programming, applied statistical analyses, and scientific communication—all from the convenience of their browser, with no installation required.\\n\\nI am working on a new Epi-Workbench (EWB) lesson titled:\\n\\n{{lesson_name}}\\n\\n## Task\\n\\nI want your help writing Code Submission Tests (CSTs) for one or more coding exercises in this lesson.\\n\\nThese CSTs will run on the Epi-Workbench platform and must follow the conventions described below exactly.\\n\\n### Execution Context\\n\\nWhen CSTs run, the test code has access to two predefined objects:\\n\\n1. \`learner_code\`: A single character string containing the learner’s entire submitted code.\\n\\n2. \`learner_env\`: An R environment in which the learner’s code has already been evaluated.\\n\\nDo not assume access to any other objects unless they are explicitly created in the hidden setup block.\\n\\n### Required Testing Pattern\\n\\nAll CSTs must follow the structure below exactly.\\n\\n#### Hidden block (for local testing only)\\n\\n\`\`\`{r, type=hide}\\n# Hidden Block for Local Testing Only\\n# -----------------------------------------------------------------------------\\n\\n# Setup the simulated learner environment\\nlearner_env \<- new.env()\\nrm(list = ls(envir = learner_env), envir = learner_env)\\n\\n# Simulated learner submissions\\ncorrect \<- \\'\[INSERT\]\\'\\nwrong_1 \<- \\'\[INSERT\]\\'\\n\\n# Set the active submission\\nlearner_code \<- correct\\n\\n# Gracefully evaluate code (prevents early error from stopping tests)\\ntry(eval(parse(text = learner_code), envir = learner_env), silent = TRUE)\\n\`\`\`\\n\\n##### Rules for simulated submissions:\\n\\n- All incorrect examples must begin with "wrong\_"\\n\\n- Use descriptive names (e.g., wrong_function_used, not wrong_1)\\n\\n- Incorrect examples should reflect realistic learner mistakes\\n\\n#### Test block (evaluated on EWB)\\n\\n\`\`\`{r, type=test}\\n# 1 - Check that \`\_\_\_\_\` was replaced with something\\ntest_that(desc = "All blanks \`\_\_\_\_\` were replaced", {\\n if (grepl("\_\_\_\_", learner_code, fixed = TRUE)) {\\n fail("It looks like your submission still contains \`\_\_\_\_\`. Please replace \`\_\_\_\_\` to complete the code.")\\n } else {\\n succeed()\\n }\\n})\\n\\n# 2 - Check ...\\ntest_that(desc = "Concise statement of the test\\'s purpose", {\\n if (condition 1) {\\n fail("Meaningful failure message")\\n } else {\\n succeed()\\n }\\n})\\n\`\`\`\\n\\n### CST Authoring Rules (Important)\\n\\n#### Test structure\\n\\n- Use multiple small \`test_that()\` blocks rather than one large test.\\n\\n- Each CST should test one narrow concept or failure mode.\\n\\n- The \`desc\` argument should be a concise statement of the test’s purpose, not a question and not the solution.\\n\\n#### Failure handling\\n\\n- Do not use \`expect\_\*()\` with \`info =\`\\n\\n- All custom feedback must be written using:\\n\\n  - \`fail()\` inside \`if\` / \`else if\` branches\\n\\n  - \`succeed()\` (or \`expect_true(TRUE)\`) at the end of a successful branch\\n \\n#### Feedback style\\n\\nCustom failure messages must:\\n\\n- Be short (ideally 1–2 lines)\\n\\n- Identify where the problem is, not the full solution\\n\\n- Be guiding rather than revealing\\n\\n- Often be phrased as questions\\n\\n- Avoid showing exact correct code (unless the issue is purely mechanical)\\n\\n- Be consistent in tone and structure across CSTs\\n\\n- Use the \`glue\` package for dynamic feedback when helpful\\n\\n#### Learner experience\\n\\n- The learner will first see: \`TEST FAILED:\` followed by the \`desc\` string\\n\\n- Assume CSTs are read by learners with limited debugging experience\\n\\n- Prefer clarity and instructional alignment over cleverness\\n\\n### What to produce\\n\\nWhen responding:\\n\\n1. Write only CST code (hidden + test blocks)\\n\\n2. Match the required structure exactly\\n\\n3. Make reasonable assumptions explicit if needed\\n\\n4. Do not explain the CSTs unless asked\\n\\nWhen ready, I will provide:\\n\\n1. The exercise prompt\\n\\n2. The starter (exercise) code scaffolding \\n\\n3. The expected solution code\\n\\n4. Any edge cases to consider'

### Issues

There are some issues with this process.

1.  In the past, we explicitly asked ChatGPT not use `testthat` expectation functions (`expect_`). I can't remember why we made that decision, but I think we should start using `testthat` expectation functions (`expect_`). Here are some reasons why:
    1.  Clarity and diagnostics: Expectations produce standardized, actionable failure messages (e.g., what was expected vs what was actually observed) that are easier for learners to interpret.
    2.  Alignment with standard practice: `testthat` is the de facto unit testing framework for R. Using its expectations makes your CSTs easier to maintain, share, and extend.
    3.  Granular feedback: You can have multiple, focused expectations within a single test block, so learners get precise feedback on each aspect of their submission.
    4.  Better reporting and tooling: `expect_*` functions integrate cleanly with `testthat` when you scale up to more exercises and analytics, and they work well with code coverage and continuous integration if you ever add it.
2.  I would like to pass the entire lesson file to ChatGPT instead of passing one coding exercise at a time. I would then like ChatGPT to locate each coding exercise and fill in the remaining parts of the coding exercise (i.e, hidden code blocks, simulated learner responses, test code blocks, and CSTs). However, I think asking ChatGPT to do all of this one pass is likely to be too complicated. Therefore, I was thinking about creating a multi-step process:
    1.  User: Add scaffolding, solution block, and instructions to the lesson Quarto file.
    2.  ChatGPT: For each exercise, write simulated learner responses.
        -   Return the simulated responses in hidden code blocks.
    3.  AI: For each simulated response, write one or more `testthat` tests using `expect_` functions.
        -   Maybe create a vector store or RAG with `testthat` documentation and EWB SOPs to help ensure consistent CSTs structure and feedback messages.
    4.  AI: For each exercise, write hints.
        -   Again, use a vector store or RAG with EWB SOPs to ensure consistent hints for learners.
3.  Adopt more of a "Unit testing" approach to the CSTs.
    -   The current arrangement of code blocks in the lesson file can take up a lot of space. This doesn't affect the learner experience, but it makes the lesson files harder for the course authors to read and maintain.
    -   We can abstract the text blocks or CSTs from the lesson file because the EWB parser looks for the CSTs in the test blocks in the lesson file.
    -   Additionally, the EWB will only be aware of code that exists in the lesson file.
    -   However, the EWB platform ignores hidden code blocks. Hidden code blocks are just for the benefit of course authors - to use for local testing. So, we could abstract them from the lesson file.
    -   I think we might be able to use a full unit testing approach like we would use for package development.
        -   R Packages, Testing Basics: https://r-pkgs.org/testing-basics.html
        -   testthat Documentation: https://testthat.r-lib.org/
    -   In other words:
        -   **`{r,type=test}` blocks must stay inside the `.qmd`** (because EWB discovers/executes CSTs there).
        -   **`{r,type=hide}` blocks are ignored by EWB** and exist only to make local authoring/testing easier.
        -   So the win is: **remove hidden blocks from the lesson files entirely** and replace them with a *project-level local testing workflow* that course authors run from the console (or a one-click script / addin). That keeps the “CSTs live in the lesson” rule intact, while making lesson files much easier to read.

### Task

Please help me think through what my next steps should be. I need to balance creating a robust and efficient long-term workflow for writing and editing lesson files - including CSTs - with the need to put lessons into production on a relatively short timeline.

If you were an expert R developer and instructional designer with deep expertise in writing testthat-based Code Submission Tests (CSTs) for educational platforms, what would you suggest for my next steps?

# 🔴 Left off here

# Save Chat History

2026-01-29: I need to pause here and move on to other tasks. I'm going to save the chat history and reload it when I come back.

```{python}
# Export chat history as a markdown file
# This version is easier to pass back to to the chat object in the future.
chat_history_path_md = (conversation_folder_path / ("multi_step_creation_chat_history_" + today + ".md"))
chat.export(chat_history_path_md, overwrite=True)

# Export chat history as an HTML file
# chat_history_path_html = (conversation_folder_path / ("multi_step_creation_chat_history_" + today + ".html"))
# chat.export(chat_history_path_html, overwrite=True)
```

## Next steps

-   This feedback is good. I need to digest it a little bit more.
- Additionally, I should probably also try to create some reusable prompts for implementing these solutions.
- You might also want to try the full 5.2 model. Actually, I passed the prompt to ChatGPT 5.2 in the browser.
- I know I want to start using expect_ functions. Maybe start with a prompt that does that.
  - Figure out the best way to write a particular CST, and then save it in SOPs.
  - Later, you can create an function, add-in, and/or VS code extension.
- Tools: For something like, "add test blocks after the solution block," write a function and then give the LLM access to the function as a tool.











# 🔴 Below not used for now.
- Just keeping in case I need it later.
- Delete when this conversation is complete.

# Read the Template Prompt

```{python}
# Raw system (developer) section of the prompt template.
system_prompt = prompt_extract_system_prompt(prompt_path)

# Raw user message section of the prompt template.
user_message_raw = prompt_extract_user_message(prompt_path)

# Inject course and lesson name into user message
# user_message = prompt_fill_placeholders(
#     text = user_message_raw,
#     replacements = {
#         "lesson_name": lesson_name,
#         "course_name": course_name,
#     },
# )
```

# Pass the user message to the AI Assistant

```{python}
chat.chat(user_message_raw, echo='none')
```

## Response

```{python}
#| echo: false
print(chat)
```

# Create a Prompt for Converting CSTs

```{python}
user_message="""
Thank you for the guidance. As part of my migration plan, I would like ChatGPT to pass existing CSTs to ChatGPT. Then, I would like ChatGPT to help me convert them to expect_* style. Will you please generate a reuseable prompt for that purpose?
"""
```

```{python}
chat.chat(user_message, echo='none')
```

## Response

```{python}
#| echo: false
print(chat)
```

# Pass the entire file

```{python}
user_message="""
It would be even better if I could just pass the entire contents of the lesson to ChatGPT in markdown format. Then, ChatGPT could locate the test blocks in it's own, and return updated CSTs for all test blocks in the lesson.
"""
```

```{python}
chat.chat(user_message, echo='none')
```

## Response

```{python}
#| echo: false
print(chat)
```

# Follow-up and Refinement

A section for optional follow-up questions and refinements. - Optionally, create a fork of the chat. `import copy; chat_fork = copy.deepcopy(chat)` - Asks follow-up questions in interactive mode - `chat.app()`.

# Conclusions 2026-01-28

-   I think there's enough there for me to create a prompt.
-   I'm not going to use all of it, but I'll use some of it.
-   Next steps:
    -   Write a prompt to upadate individual CSTs.
    -   Write a prompt to update all existing test blocks in a lesson file.
    -   I also need to add additional COL code submission questions to the Descriptive Analysis I lessons.

# Save Chat History

Optional, but necessary to return to the conversation later.

```{python}
# Export chat history as a markdown file
# This version is easier to pass back to to the chat object in the future.
chat_history_path_md = (conversation_folder_path / ("Use_expect_chat_history_" + today + ".md"))
chat.export(chat_history_path_md, overwrite=True)

# Export chat history as an HTML file
chat_history_path_html = (conversation_folder_path / ("Use_expect_chat_history_" + today + ".html"))
chat.export(chat_history_path_html, overwrite=True)

# Export chat history as a json file
# Useful for saving/loading the history to/from a database or file.
# chat_history_path_json = (lesson_folder_path / "ai" / ("lesson_revisions_history_" + today + ".json"))
# with open(chat_history_path_json, "w", encoding="utf-8") as f:
#     json.dump(
#         [json.loads(x) for x in turns_json],
#         f,
#         indent=2,
#         ensure_ascii=False
#     )
```

## Condense

You can also optionally condense the history by asking the LLM to summarize it. This is useful if you want to keep the context of the conversation, but don’t want to provide the entire history to the model on each new submission.

```{python}
# chat.chat("Can you summarize our conversation so far?")
# chat.set_turns([chat.get_last_turn()])
```

## Load chat history

To optionally load the chat history later.

```{python}
# Import chat history as a md file
# chat_history_path_md = (lesson_folder_path / "ai" / ("lesson_revisions_history_" + "yyyy-mm-dd" + ".md"))
# history_md = chat_history_path_md.read_text(encoding="utf-8")

# Import chat history as a json file
# chat_history_path_json = (lesson_folder_path / "ai" / ("lesson_revisions_history_" + "yyyy-mm-dd" + ".json"))
# history_json = json.loads(chat_history_path_json.read_text(encoding="utf-8"))
```

# Render to HTML

Render to HTML (Optional) - Click the Preview button. - Type `quarto render "path/Lesson Revisions.qmd"` in the terminal.