---
title: "Chat Template"
id: chat_template
author: brad_cannell
date: "2025-12-23"
date-modified: "2025-12-23"
audience: course_authors
category: course_development
tags:
  - tag1
  - tag2
format:
  html:
    toc: true
    embed-resources: true
execute:
  eval: false
---

<!--
YAML HEADER INSTRUCTIONS

Complete the YAML header above for every prompt. These fields support organization, reuse, and future search, but should remain lightweight and human-readable.

FIELD DESCRIPTIONS
- title: A brief, human-readable description of what the prompt does. Written in Title Case.
- id: Use a short, stable, human-readable identifier in lowercase snake_case.
- task: The primary type of work the prompt performs (e.g., text_revision, code_review). Written in snake_case.
- audience: The intended audience for the output (e.g., academic, student, developer). Written in snake_case.
- author: Name of the prompt author. Written in snake_case.
- created: Date the prompt was first created (YYYY-MM-DD).
- last_updated: Date of the most recent substantive revision (YYYY-MM-DD).
- project: Associated project, course, or initiative, or null if broadly reusable. Written in snake_case.
- domain: Subject area or disciplinary context (e.g., epidemiology, data_science, general). Written in snake_case.
- tags: Optional keywords to support browsing and future search. Written in snake_case.

ID GUIDELINES
- id: Use a short, stable, human-readable identifier in lowercase snake_case.
- Base the id on the prompt’s core purpose using a verb–object pattern
  (e.g., rewrite_academic_tone, review_r_code_style).
- Avoid dates, version numbers, model names, project names, or author names.
- Treat the id as permanent; change it only if the prompt’s fundamental purpose changes.

Avoid including model names or API parameters in the YAML header.
-->

# Overview

This section is for the page author or future maintainers, not the AI assistant.

Use this space to capture motivation, background, assumptions, or reminders that should not be included in the system prompt or user instructions.

```{r}
#| echo: false

library(ellmer)
```

```{r}
#| echo: false

# Load strip_yaml_header()
source('/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/scripts/strip_yaml_header.R')

# Load prompt_extract_system_prompt()
source('/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/scripts/prompt_extract_system_prompt.R')

# Load prompt_extract_user_message()
source('/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/scripts/prompt_extract_user_message.R')

# Load prompt_fill_placeholders()
source('/Users/bradcannell/Library/CloudStorage/GoogleDrive-brad.cannell@epi-workbench.com/My Drive/AI Assistant/scripts/prompt_fill_placeholders.R')
```

# Retrieve the API key and Expose it to ellmer

```{r}
Sys.setenv(OPENAI_API_KEY = keyring::key_get("ewb-general"))
Sys.setenv(GOOGLE_API_KEY = keyring::key_get("ewb-general-gemini"))
Sys.setenv(ANTHROPIC_API_KEY = keyring::key_get("ewb-general-claude"))
```

# Read the Template Prompt

```{r}
system_prompt <- prompt_extract_system_prompt('path to prompt')
```

```{r}
user_message_raw <- prompt_extract_user_message('path to prompt')
```

# Create Chat Object

Periodically check to make sure you are running the best model for the task at hand.

```{r}
chat <- chat_openai(
  system_prompt = system_prompt,
  model = "gpt-5-nano-2025-08-07",
  echo = "none"
)
```

# Prompt

```{r}
conversation <- chat$chat(user_message_raw)
```

## Response

```{r}
#| echo: false
cat(conversation)
```

````markdown
Copy and paste response here.
````

# Write the response to a markdown file for review

Optional

```{r}
writeLines(conversation, "template_conversation.md")
```

# Save Chat Object

Optional, but necessary to return to the conversation later.

```{r}
saveRDS(conversation, "template_conversation_date.rds")
```

Render to HTML