---
title: "Categories and Tags"
author: "Brad Cannell"
date: "2025-12-23"
date-modified: "2025-12-22"
category: "Course Development"
tags: NULL
format:
  html:
    embed-resources: true
execute:
  eval: false
---

# Overview

This wiki page is to help course authors appropriately use request and assign tags.
- **2025-12-23:** 
  - I downloaded an existing ClickUp Doc draft of this SOP (ClickUp-Categories and Tags-20251223110159.md). 
  - I made some adjustments to the ClickUp draft to create an initial draft SOP to pass to the AI assistant.
  - My goal today is to create a "good enough" set of guidelines for using tags. The tagging system isn't finalized yet. We are still waiting on:
    - [Tags: Course Authors View Tag List](https://trello.com/c/7GLitlrp/114-tags-course-authors-view-tag-list)
    - [Tags: Type to Search for Tags](https://trello.com/c/5LnNpNdV/115-tags-type-to-search-for-tags)
    - [Tags: Tag Lessons Instead of Courses](https://trello.com/c/mCQyAXYf/116-tags-tag-lessons-instead-of-courses)
    - [Wiki: Add Images to Course Author Wiki Pages](https://trello.com/c/JRvkjSdH/117-add-images-to-course-author-wiki-pages)

# Load Packages

```{r}
library(ellmer)
```

# Retrieve the API key and Expose it to ellmer

```{r}
Sys.setenv(GOOGLE_API_KEY = keyring::key_get("ewb-wiki"))
Sys.setenv(ANTHROPIC_API_KEY = keyring::key_get("test-claude-brad"))
Sys.setenv(OPENAI_API_KEY = keyring::key_get("ewb-wiki-openai"))
```

# Initial Draft

2025-12-23: 
  - I downloaded an existing ClickUp Doc draft of this SOP (ClickUp-Categories and Tags-20251223110159.md). 
  - I made some adjustments to the ClickUp draft to create an initial draft SOP to pass to the AI assistant.
  - My goal today is to create a "good enough" set of guidelines for using tags. The tagging system isn't finalized yet.
  
## System prompt
  
```{r}
system_prompt <- paste(
  "You are a technical writer with expertise in writing standard operating procedures (SOPs),",
  "guidelines, knowledge bases, and wikis for organizations.",
  "Your task is to revise the provided draft wiki page for clarity, accuracy, and usability.",
  "Use the provided context for background understanding only.",
  "Do NOT rewrite or summarize the background context unless explicitly instructed.",
  "Write in a clear, professional, instructional tone appropriate for course authors.",
  sep = " "
)
```

## Create chat object

```{r}
chat <- chat_google_gemini(
  system_prompt = system_prompt,
  model = "gemini-2.5-pro",
  echo = "none"
)
```

## Create user message

### Context

```{r}
# Read About EWB (raw, including YAML if present)
about_ewb_raw <- readLines(here::here('about_ewb.md'), warn = FALSE)

# Read About Me (raw, including YAML if present)
about_me_raw <- readLines(here::here('about_me.md'), warn = FALSE)

# Read the draft wiki page (this file does NOT have YAML)
draft_raw <- readLines(here::here('conversations', 'Wiki Page - Tags', 'Initial Draft.md'), warn = FALSE)
```

#### Function for removing the yaml header

```{r}
# This function removes a YAML header if it exists at the top of a markdown file.
# A YAML header is assumed to:
#   - start with '---' on the first line
#   - end with the next '---'
strip_yaml_header <- function(lines) {
  
  # Check whether the first non-empty line is a YAML start marker
  first_non_empty <- which(nzchar(trimws(lines)))[1]
  
  if (!is.na(first_non_empty) && trimws(lines[first_non_empty]) == "---") {
    
    # Find the line number of the closing YAML delimiter
    yaml_end <- which(trimws(lines) == "---")[-1][1]
    
    # Return everything *after* the YAML block
    return(lines[(yaml_end + 1):length(lines)])
    
  } else {
    
    # If no YAML header is present, return the original lines unchanged
    return(lines)
  }
}

# For testing
# strip_yaml_header(about_me)
```

Strip YAML from context files

```{r}
about_ewb_clean <- strip_yaml_header(about_ewb_raw)
about_me_clean  <- strip_yaml_header(about_me_raw)
```

Collapse each section into a single string

```{r}
about_ewb_text <- paste(about_ewb_clean, collapse = "\n")
about_me_text  <- paste(about_me_clean, collapse = "\n")
draft_text     <- paste(draft_raw, collapse = "\n")
```

Combine into one structured prompt string

```{r}
# Build the final prompt with explicit section labels
prompt_text <- paste(
  "## Context: About Epi-Workbench (EWB)\n",
  about_ewb_text,
  "\n\n---\n\n",
  "## Context: About the Author\n",
  about_me_text,
  "\n\n---\n\n",
  "## Draft Wiki Page: Categories and Tags\n",
  draft_text,
  sep = ""
)
```

Checkpoint — inspect the prompt

```{r}
# Print the first ~80 lines to verify structure and labeling
# cat("===== BEGIN PROMPT PREVIEW =====\n")
# cat(paste(head(strsplit(prompt_text, "\n")[[1]], 80), collapse = "\n"))
# cat("\n===== END PROMPT PREVIEW =====\n")
```

Create a combined user message

```{r}
# --------------------------------------------
# Step 2: Define the user message
# --------------------------------------------

# The user message contains the actual material to work with.
# This wraps the structured prompt string created earlier.
user_message <- paste(
  "Please revise the following wiki page based on the provided context.\n\n",
  prompt_text,
  sep = ""
)
```

## Send the request to the API

```{r}
response <- chat$chat(user_message)
```

## Response

Write the response to a markdown file for review.

```{r}
writeLines(response, "gemini_draft_01.md")
```
